Which ITG Stepcharts are Turniest?

### 2017-03-16 itg/2/paper.tex
finish the paper

It is tempting to assign the facings $L$, $U$, $U$, and $R$ respectively to the $LL$, $DD$, $UU$, and $RR$ footswitches.
However, Figure~\ref{fig:fs}(c) shows that if a footswitch begins with a crossover on $U$, the facing should be reversed:
the $RR$ footing should face $L$, and $LL$ should face $R$.
``Spin-switches'' with $D$ facing are also theoretically possible,
arising from patterns such as $LURDDL$ or $LDRUUL$, although I know of no real-world chart which does this (hmm, tempting...).

@@

To fairly represent a human player's desire to step in the least turny of ambiguous ways,
I extended the algorithm to provide either the assigned facing, from above, or its polar opposite,
chosen at runtime by whichever is closer to the presequent facing.
This restores the maximum overall chart turniness to $\mathcal{T}=2$,
a new example of which is shown in Figure~\ref{fig:fs-turniness}(c).

@@

Another approach could assign such a footswitch the opposite footing of the previous facing,
regardless of the arrow itself;
so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
Anyway, remember that all-candle switch stream in (c). We'll see it again soon.

@@

@@

\begin{figure*}[p]

@@

\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\

@@

\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\

@@

\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\

@@

\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\

@@

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{paradise-lost-false-positive.png}
		\\
		(a) Paradise Lost \\
		(16, Cirque du Lykan)
	\end{tabular}
	%\begin{tabular}{c}
	%	\includegraphics[width=0.18\textwidth]{false-positive-toromi-hearts.png}
	%	\\
	%	(a) Toromi Hearts 2 \\
	%	(18, Cirque du Zonda)
	%\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.18\textwidth]{hearbeat-rollstream-doublestep.png}
		\\
		(b) Heartbeat \\
		(13, TranceNation)
	\end{tabular}
	\end{center}
	\caption{The doublesteps in some streamy charts must be identified, and the stream split, lest ``too much'' of the following stream appear completely crossed-over.}
	\label{fig:force-doublestep}
\end{figure}

\newcommand\Stream{\ensuremath{\mathcal{S}}}
\newcommand\Doublestep{\ensuremath{\mathcal{DS}}}
\newcommand\nflip{\ensuremath{n_{\mathfrak{flip}}}}
\newcommand\percentflip{\ensuremath{\%_{\mathfrak{flip}}}}
\begin{algorithm}[h!]
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Invariant}
	\Input{\Stream, a step sequence $s_0 \dots s_n$}
	\Output{$\forall s_i, s_j \in \Stream, j=i+1 \rightarrow \mathsf{\neg StreamBoundary}(s_i,s_j)$}
	\Input{\nflip, heuristic minimum length} %of flipped stream to doublestep into}
	\Input{$\percentflip$, heuristic percentage, initially 100}
	\For{$i \in \mathsf{length}(\Stream) \wedge \neg {\sf defined}(i_{\Doublestep})$}{
		$\Stream' \leftarrow \{ s_k | s_k \in \mathsf{LRs}(\{ s_j | s_j \in \Stream \wedge j \ge i \}) \wedge k \le \nflip \}$ \\
		\If{$|\Stream'| = \nflip$ $\wedge$ $|\mathsf{Crossovers}(\Stream')| \ge \percentflip \times |\Stream'|$}{
			$i_{\Doublestep} \leftarrow i$
		}
	}
	\eIf{{\sf defined}($i_{\Doublestep})$}{
		\If{$i_{\Doublestep} = 0$}{
			% gotta write "tail stream" as complicatedly as possible
			$i_{\Doublestep} \leftarrow \mathsf{FindUnflippedSection}(\{ s_i | s_i \in \Stream \wedge i \ne 0 \})$
			%\mathsf{tail}(\Stream))$
		}
		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i < i_{\Doublestep} \}$)
		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i \ge i_{\Doublestep} \}$)
	}{
		{\sf CommitStream}(\Stream)
	}
	\caption{{\sf HeuristicallyDoublestep}(\Stream)}
	\label{alg:heuristic}
\end{algorithm}

@@

\section{Experimental Results}

% TODO: Put a * on charts where you're going to put a pict of mario man
% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended

% TODO: Make a special mention of sandstorm and put a pict of mario man
% TODO: put the above in an honourable mention subsection, and also mention the slow train and sidefoots

% Stuff to talk about.
% Dataset size now >10k
% I reran turniness on fancy footwork and it actually got turnier (double check dis), b/c of all the candles
% 9-15 is clearly the sweet spot for techni jawnz

% By author/by pack: I excluded 1s-3s, admittedly biased against konami, on the pretense of excluding nonsense false positives, but even after that, konami is still the crossover king
% Step authors are limited to 10 or more charts

% Miscellaneous charts I want to exhibit:
% Conflict forced doublesteps
% Heart shooter 'sidefoots'
% Boom Bombay - triple xover jack footed as sideswitch
% Toxic - jack+doublestep footed as sideswitch + xover stream instead
% H'ble mention to sandstorm for jacks

@@

		 7 & Tetris                   & CuoReNeRo             &  58 \\
		 8 & Pulse                    & CuoReNeRo             &  94 \\
		 9 & MAX Forever              & CuoReNeRo             & 107 \\
		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo             & 141 \\

@@

@@

		\multicolumn{4}{c}{\em no 16s-19s with $\ge$10\% XOs} \\

@@

@@

		 8 & Pulse            & CuoReNeRo             &  41 \\

@@

@@

	\caption{Charts with the highest percentage of FSs among total steps.}

@@

@@

		12 & Beautiful Synergy & Keyboard Collab. III & 14 \\

@@

	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}

@@

% authors
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|r|r|l}
		\bf Author & \bf Charts & \bf Total Steps & \bf FS\% \\
		\hline
		Konami        & 530 & 114623 & 8.03 \\
		sssmsm        &  41 &  17219 & 6.55 \\
		NEMORIGINAL   &  44 &  20165 & 5.36 \\
		M. Emirzian   &  23 &   9307 & 5.16 \\
		J. DeGarmo    &  16 &   4362 & 4.86 \\
		D. Renzetti   &  18 &   7877 & 4.47 \\
		R. McKanna    &  47 &  14855 & 4.23 \\
		M. Puls       &  26 &   8379 & 4.18 \\
		King of Light &  24 &   8817 & 4.13 \\
		D. Bernardone & 217 &  76290 & 4.10 \\
		D. D'Amato    & 107 &  36265 & 3.92 \\
		bblum         &  32 &  32382 & 3.76 \\
		\multicolumn{4}{c}{\normalsize\dots} \\
		B. Vergara    &  13 &  15454 & 0.091 \\
		Aoreo         &  21 &  27990 & 0.089 \\
		Zaia          & 368 & 448460 & 0.080 \\
		t0ni          &  85 & 128964 & 0.058 \\
		Burn          &  27 &  60052 & 0.057 \\
		Dirk          &  12 &  21996 & 0.055 \\
		Happy Feet    &  30 &  57888 & 0.040 \\
		@@            &  63 & 199530 & 0.026 \\
		Arvin         &  79 & 108612 & 0.023 \\
		Drazu         & 153 & 221460 & 0.021 \\
		teejusb       &  11 &  11298 & 0.018 \\
		Fraxtil       &  19 &  25612 & 0 \\
	\end{tabular}
	\end{center}
	\caption{Chart authors with the highest/lowest XO\%.}
\end{table}
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|r|r|l}
		\bf Author & \bf Charts & \bf Total Steps & \bf FS\% \\
		\hline
		Konami        & 530 & 114623 & 2.29 \\
		bblum         &  32 &  32382 & 1.29 \\
		M. Puls       &  26 &   8379 & 1.16 \\
		R. McKanna    &  47 &  14855 & 1.11 \\
		mudkyp        &  63 &  42792 & 1.09 \\
		S. Venkat     &  24 &  11084 & 0.96 \\
		K. Ward       & 281 &  86475 & 0.89 \\
		xRGTMx        &  19 &  11734 & 0.88 \\
		sssmsm        &  41 &  17219 & 0.85 \\
		D. Bernardone & 217 &  76290 & 0.83 \\
		ATB           &  31 &  20673 & 0.82 \\
		Happy Feet    &  30 &  57888 & 0.82 \\
		\multicolumn{4}{c}{\normalsize\dots} \\
		Hsarus        &  18 &  70162 & 0.070 \\
		Drazu         & 153 & 221460 & 0.057 \\
		@@            &  63 & 199530 & 0.037 \\
		Revolver      &  11 &   8302 & 0.036 \\
		T. Swag       &  13 &  22909 & 0.031 \\
		Dirk          &  12 &  21996 & 0.023 \\
		B. Vergara    &  13 &  15454 & 0.019 \\
		warpdr!ve     &  16 &  19090 & 0.016 \\
		Burn          &  27 &  60052 & 0.015 \\
		teejusb       &  11 &  11298 & 0.009 \\
		t0ni          &  85 & 128964 & 0.002 \\
		S. Tofu       &  26 &  34609 & 0
	\end{tabular}
	\end{center}
	\caption{Chart authors with the highest/lowest FS\%.}
\end{table}
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|r|r|l}
		\bf Author & \bf Charts & \bf Total Steps & \bf JK\% \\
		\hline
		King of Light &  24 &   8817 & 12.94 \\
		R. McKanna    &  47 &  14855 & 9.95 \\
		Konami        & 530 & 114623 & 9.63 \\
		P. Shanklin   &  21 &   9834 & 8.91 \\
		M. Puls       &  26 &   8379 & 8.90 \\
		K. Ward       & 281 &  86475 & 8.80 \\
		ATB           &  31 &  20673 & 8.46 \\
		J. DeGarmo    &  16 &   4362 & 8.30 \\
		D. Bernardone & 217 &  76290 & 7.98 \\
		Renard        &  45 &  15035 & 7.94 \\
		C. Foy        & 133 &  52418 & 7.72 \\
		Yoko          &  10 &   4128 & 7.63 \\
		\multicolumn{4}{c}{\normalsize\dots} \\
		bblum         &  32 &  32382 & 1.56 \\
		B. Vergara    &  13 &  15454 & 1.31 \\
		Arvin         &  79 & 108612 & 1.30 \\
		Drazu         & 153 & 221460 & 1.06 \\
		Zaia          & 368 & 448460 & 0.86 \\
		Aoreo         &  21 &  27990 & 0.78 \\
		T. Swag       &  13 &  22909 & 0.70 \\
		@@            &  63 & 199530 & 0.61 \\
		warpdrive     &  16 &  19090 & 0.49 \\
		Dirk          &  12 &  21996 & 0.35 \\
		Burn          &  27 &  60052 & 0.28 \\
		Hsarus        &  18 &  70162 & 0.27 \\
	\end{tabular}
	\end{center}
	\caption{Chart authors with the highest/lowest JK\%.}
\end{table}

@@

@@

@@

@@

@@

@@

Let's be honest: this isn't gonna be a trilogy.

With that out of the way, here are some things that would be cool to implement in a fantasy universe with infinite free time.

% TODO
% forced doublesteps via holds
% - bracketjump footswitches - matador from valex 8
% understanding mine cues for doublesteps

>>>>

It is tempting to assign the facings $L$, $U$, $U$, and $R$ respectively to the $LL$, $DD$, $UU$, and $RR$ footswitches.
However, Figure~\ref{fig:fs}(c) shows that if a footswitch begins with a crossover on $U$, the facing should be reversed:
the $RR$ footing should face $L$, and $LL$ should face $R$.
``Spin-switches'' with $D$ facing are also theoretically possible,
arising from patterns such as $LURDDL$ or $LDRUUL$,
%although I know of no real-world chart which does this (hmm, tempting...).
or similarly, ``270-switches'', as shown in Figure~\ref{fig:real-world-turniness}(a).

@@

To fairly represent a human player's desire to step in the least turny of ambiguous ways,
I extended the algorithm to provide either the assigned facing, from above, or its polar opposite,
chosen at runtime by whichever is closer to the presequent facing.
This restores the maximum overall chart turniness to $\mathcal{T}=2$,
a new example of which is shown in Figure~\ref{fig:fs-turniness}(c).
Figure~\ref{fig:real-world-turniness}(b) also shows a real-world chart exhibiting this pattern.

@@

Another approach could assign such a footswitch the opposite footing of the previous facing,
regardless of the arrow itself;
so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
%Anyway, remember that all-candle switch stream in (c). We'll see it again soon.

@@

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{web33-dd-switch-270.png}
		\\
		(a) Web33,260.8 \\
		(12, Rikame 5)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{fuego-candle-fs.png}
		\\
		(b) Fuego \\
		(12, Best of Gazebo)
	\end{tabular}
	\end{center}
	\caption{Real-world examples of turny footswitches.}
	\label{fig:real-world-turniness}
\end{figure}

@@

\begin{figure*}[ht]

@@

%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- reverse the stream's footing if more L/R steps were crossed over than not.}} \\

@@

%\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
%\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
%\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\

@@

%\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
%\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
%\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
%\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\

@@

%\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\

@@

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{paradise-lost-false-positive.png}
		\\
		(a) Paradise Lost \\
		(16, Cirque du Lykan)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{hearbeat-rollstream-doublestep.png}
		\\
		(b) Heartbeat \\
		(13, TranceNation)
	\end{tabular}
	\end{center}
	\caption{The doublesteps in some streamy charts must be identified, and the stream split, lest ``too much'' of the following stream appear completely crossed-over.}
	\label{fig:force-doublestep}
\end{figure}

\newcommand\Stream{\ensuremath{\mathcal{S}}}
\newcommand\Doublestep{\ensuremath{\mathcal{DS}}}
\newcommand\nflip{\ensuremath{n_{\mathfrak{flip}}}}
\newcommand\percentflip{\ensuremath{\%_{\mathfrak{flip}}}}
\begin{algorithm}[h]
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Invariant}
	\Input{\Stream, a step sequence $s_0 \dots s_n$}
	\Output{$\forall s_i, s_j \in \Stream, j=i+1 \rightarrow \mathsf{\neg StreamBoundary}(s_i,s_j)$}
	\Input{\nflip, heuristic minimum length} %of flipped stream to doublestep into}
	\Input{$\percentflip$, heuristic percentage, initially 100}
	\For{$i \in \mathsf{length}(\Stream) \wedge \neg {\sf defined}(i_{\Doublestep})$}{
		$\Stream' \leftarrow \{ s_k | s_k \in \mathsf{LRs}(\{ s_j | s_j \in \Stream \wedge j \ge i \}) \wedge k \le \nflip \}$ \\
		\If{$|\Stream'| = \nflip$ $\wedge$ $|\mathsf{Crossovers}(\Stream')| \ge \percentflip \times |\Stream'|$}{
			$i_{\Doublestep} \leftarrow i$
		}
	}
	\eIf{{\sf defined}($i_{\Doublestep})$}{
		\If{$i_{\Doublestep} = 0$}{
			% gotta write "tail stream" as complicatedly as possible
			$i_{\Doublestep} \leftarrow \mathsf{FindUnflippedSection}(\{ s_i | s_i \in \Stream \wedge i \ne 0 \})$
			%\mathsf{tail}(\Stream))$
		}
		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i < i_{\Doublestep} \}$)
		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i \ge i_{\Doublestep} \}$)
	}{
		{\sf CommitStream}(\Stream)
	}
	\caption{{\sf HeuristicallyDoublestep}(\Stream)}
	\label{alg:heuristic}
\end{algorithm}

@@

\section{Evaluation}

Our experimental corpus has grown considerably since last year, and now comprises 11,666 stepcharts.
I ran the crossoveriness/etcetera algorithm on all of them,
and counted the total steps (not including jumps), crossovers, footswitches, jacks, forced doublesteps, and crossover switches for each.
I also grouped the charts by author and by song pack to calculate each author's/pack's overall crossoveriness/etc.
You can view the entire dataset at \url{http://tinyurl.com/crossoveriest}.

Tables
\ref{tab:chart-xo},
\ref{tab:chart-xop},
\ref{tab:chart-fs},
\ref{tab:chart-fsp},
\ref{tab:chart-xf},
\ref{tab:author-xo},
\ref{tab:author-fs},
\ref{tab:author-jk},
\ref{tab:pack-xo},
\ref{tab:pack-fs},
and
\ref{tab:pack-jk}
summarize the dataset as ``leaderboards'' for each category.
They present the data the same way as last year \cite{turniness},
so I won't explain it again.
Suffice to say that ITG enthusiasts should use their personal chart style preferences to navigate these tables and find song or pack recommendations or feel proud of themselves or whatever.

In the by-author analysis, I excluded authors with fewer than 10 charts, like last time.
Also, in by-author and by-pack, I excluded 1-, 2-, and 3-foot charts,
on the pretense that they often ignore the alternating feet assumption
(though I also admit this biases the analysis against DDR/Konami).
%(though I also admit some bias in wanting to unseat DDR/Konami as the crossover king).
Nevertheless, DDR charts generally ruled the single-digits,
even in footswitchiness,
showing perhaps more technical depth than I gave them credit for.
%Also, congratulations to Fraxtil and S. Tofu, who managed to write not a single crossover or footswitch, respectively.

By the way, the theoretical maxima for XO\%, FS\%, and JK\% are 50-$\epsilon$, 100-$\epsilon$, and 100-$\epsilon$, respectively \cite{epsilon}.

@@

		 7 & Tetris                   & CuoReNeRo M'PacK      &  58 \\
		 8 & Pulse                    & CuoReNeRo M'PacK      &  94 \\
		 9 & MAX Forever              & CuoReNeRo M'PacK      & 107 \\
		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo M'PacK      & 141 \\

@@

	\label{tab:chart-xo}

@@

		\multicolumn{4}{c}{\em no 16s-19s with $\ge$10\% XO} \\

@@

	\label{tab:chart-xop}

@@

		 8 & Pulse            & CuoReNeRo M'PacK      &  41 \\

@@

	\label{tab:chart-fs}

@@

	\caption{Charts with the highest percentage of FSs.} %among total steps.}
	\label{tab:chart-fsp}

@@

\begin{table}[h!]

@@

		12 & Web 33,260.8      & Rikame's Simfiles 5        & 16 \\

@@

	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off to exclude a bunch of ambiguously-patterned charts from DDR.)}
	%because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}
	\label{tab:chart-xf}

@@

% authors
\begin{table}[p]
	\begin{center}
		\small
	\begin{tabular}{l|r|r|l}
		\bf Author & \bf Charts & \bf Total Steps & \bf FS\% \\
		\hline
		Konami        & 530 & 114623 & 8.03 \\
		sssmsm        &  41 &  17219 & 6.55 \\
		NEMORIGINAL   &  44 &  20165 & 5.36 \\
		M. Emirzian   &  23 &   9307 & 5.16 \\
		J. DeGarmo    &  16 &   4362 & 4.86 \\
		D. Renzetti   &  18 &   7877 & 4.47 \\
		R. McKanna    &  47 &  14855 & 4.23 \\
		M. Puls       &  26 &   8379 & 4.18 \\
		King of Light &  24 &   8817 & 4.13 \\
		D. Bernardone & 217 &  76290 & 4.10 \\
		D. D'Amato    & 107 &  36265 & 3.92 \\
		bblum         &  32 &  32382 & 3.76 \\
		\multicolumn{4}{c}{\normalsize\dots} \\
		B. Vergara    &  13 &  15454 & 0.091 \\
		Aoreo         &  21 &  27990 & 0.089 \\
		Zaia          & 368 & 448460 & 0.080 \\
		t0ni          &  85 & 128964 & 0.058 \\
		Burn          &  27 &  60052 & 0.057 \\
		Dirk          &  12 &  21996 & 0.055 \\
		Happy Feet    &  30 &  57888 & 0.040 \\
		@@            &  63 & 199530 & 0.026 \\
		Arvin         &  79 & 108612 & 0.023 \\
		Drazu         & 153 & 221460 & 0.021 \\
		teejusb       &  11 &  11298 & 0.018 \\
		Fraxtil       &  19 &  25612 & 0 \\
	\end{tabular}
	\end{center}
	\caption{Chart authors with the highest/lowest XO\%.}
	\label{tab:author-xo}
\end{table}
\begin{table}[p]
	\begin{center}
		\small
	\begin{tabular}{l|r|r|l}
		\bf Author & \bf Charts & \bf Total Steps & \bf FS\% \\
		\hline
		Konami        & 530 & 114623 & 2.29 \\
		bblum         &  32 &  32382 & 1.29 \\
		M. Puls       &  26 &   8379 & 1.16 \\
		R. McKanna    &  47 &  14855 & 1.11 \\
		mudkyp        &  63 &  42792 & 1.09 \\
		S. Venkat     &  24 &  11084 & 0.96 \\
		K. Ward       & 281 &  86475 & 0.89 \\
		xRGTMx        &  19 &  11734 & 0.88 \\
		sssmsm        &  41 &  17219 & 0.85 \\
		D. Bernardone & 217 &  76290 & 0.83 \\
		ATB           &  31 &  20673 & 0.82 \\
		Happy Feet    &  30 &  57888 & 0.82 \\
		\multicolumn{4}{c}{\normalsize\dots} \\
		Hsarus        &  18 &  70162 & 0.070 \\
		Drazu         & 153 & 221460 & 0.057 \\
		@@            &  63 & 199530 & 0.037 \\
		Revolver      &  11 &   8302 & 0.036 \\
		T. Swag       &  13 &  22909 & 0.031 \\
		Dirk          &  12 &  21996 & 0.023 \\
		B. Vergara    &  13 &  15454 & 0.019 \\
		warpdr!ve     &  16 &  19090 & 0.016 \\
		Burn          &  27 &  60052 & 0.015 \\
		teejusb       &  11 &  11298 & 0.009 \\
		t0ni          &  85 & 128964 & 0.002 \\
		S. Tofu       &  26 &  34609 & 0
	\end{tabular}
	\end{center}
	\caption{Chart authors with the highest/lowest FS\%.}
	\label{tab:author-fs}
\end{table}
\begin{table}[p]
	\begin{center}
		\small
	\begin{tabular}{l|r|r|l}
		\bf Author & \bf Charts & \bf Total Steps & \bf JK\% \\
		\hline
		King of Light &  24 &   8817 & 12.94 \\
		R. McKanna    &  47 &  14855 & 9.95 \\
		Konami        & 530 & 114623 & 9.63 \\
		P. Shanklin   &  21 &   9834 & 8.91 \\
		M. Puls       &  26 &   8379 & 8.90 \\
		K. Ward       & 281 &  86475 & 8.80 \\
		ATB           &  31 &  20673 & 8.46 \\
		J. DeGarmo    &  16 &   4362 & 8.30 \\
		D. Bernardone & 217 &  76290 & 7.98 \\
		Renard        &  45 &  15035 & 7.94 \\
		C. Foy        & 133 &  52418 & 7.72 \\
		Yoko          &  10 &   4128 & 7.63 \\
		\multicolumn{4}{c}{\normalsize\dots} \\
		bblum         &  32 &  32382 & 1.56 \\
		B. Vergara    &  13 &  15454 & 1.31 \\
		Arvin         &  79 & 108612 & 1.30 \\
		Drazu         & 153 & 221460 & 1.06 \\
		Zaia          & 368 & 448460 & 0.86 \\
		Aoreo         &  21 &  27990 & 0.78 \\
		T. Swag       &  13 &  22909 & 0.70 \\
		@@            &  63 & 199530 & 0.61 \\
		warpdrive     &  16 &  19090 & 0.49 \\
		Dirk          &  12 &  21996 & 0.35 \\
		Burn          &  27 &  60052 & 0.28 \\
		Hsarus        &  18 &  70162 & 0.27 \\
	\end{tabular}
	\end{center}
	\caption{Chart authors with the highest/lowest JK\%.}
	\label{tab:author-jk}
\end{table}

@@

\begin{table}[p]

@@

	\label{tab:pack-xo}

@@

	\label{tab:pack-fs}

@@

	\label{tab:pack-jk}

@@

\section{Discussion}

To verify the algorithm's accuracy,
I manually inspected a random (read: not random) sample of the charts at or near the top of the various leaderboards
(read: I played a lot of Stepmania).
I also consulted a leading expert in the field of automated ITG chart analysis (read: myself),
who reported that the algorithm is infinitely more accurate than the prior state-of-the-art.

Honestly though, it works really well.
Last year's algorithm was often finicky and prone to all sorts of false-positives,
while this one plays ITG in a recognizably human way (almost) without fail.
It was a joy to use.
%Certain rare patterns could still confuse it, while on others...

{\bf Surprises.}
On occasion, the algorithm surprised me by stepping with crossover footswitches which,
at first glance, I would probably jack or double-step.
However, these always proved to be perfectly valid alternative footings,
in some cases requiring considerable look-ahead.
In Figure~\ref{fig:surprise-sideswitch}(a),
the chart repeats $L$ (later $R$) thrice,
beginning with the right (later, left) foot.
While a human player would jack these repeated arrows,
the crossoveriness algorithm performs a double-footswitch,
effectively reducing the total crossover steps by 1 each time.
In Figure~\ref{fig:surprise-sideswitch}(b),
a mine cues the player to double-step with her right foot,
but the crossoveriness algorithm can begin this section already crossed-over,
owing to an earlier $L$ jack on which it could switch feet.

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{boom-bombay-double-sideswitch-instead-of-jack.png}
		\\
		(a) Dr. Boom-Bombay \\
		(9, fort rapids vii)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{toxic-xoverfs-avoid-doublestep.png}
		\\
		(b) Toxic \\
		(10/13, Sex'y Violation 2)
	\end{tabular}
	\end{center}
	\caption{Sometimes the algorithm was smarter than me.}
	\label{fig:surprise-sideswitch}
\end{figure}

{\bf Honourable Mentions.}
I omitted a table for the jackiest charts,
on account of most of them being either trivial beginner charts or extra-long megamixes.
One deserves a special mention:
Sandstorm (Jimmy Jawnz 2), shown in Figure~\ref{fig:hble-mention}(a),
has more than twice as many total jacks as the next jackiest chart,
clocking in at 1049 (78.5\%) with its 15 and 992 (69.6\%) with its 17.
And looking at that chart, can't you just hear Sandstorm playing in your head already?

I also wanted to highlight the chart with the most crossover switches,
shown in Figure~\ref{fig:hble-mention}(b),
mostly because the skittle notes should add some nice variety of colour to the paper
(with apologies to the dead-tree SIGBOVIK audience reading in greyscale).
Figure~\ref{fig:hble-mention}(c), with 2nd place in crossover switches following (b),
comes with an edit chart titled ``no sidefoots'',
and to be perfectly honest I just kept saying the word ``sidefoots'' to myself and giggling a lot while writing this paper.

\begin{figure*}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{sandstorm-jacks.png}
		\\
		(a) Sandstorm \\
		(15/17, J. Jawnz II)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{heart-shooter-xover-fs.png}
		\\
		(b) Heart Shooter \\
		(11, VocaJawnz)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{slow-train-sidefoots.png}
		\\
		(c) '76 (Slow Train) \\
		(11, Mute Sims X)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{conflict-cued-doublesteps.png}
		\\
		(d) Conflict \\
		(12, Stephcharts)
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{matador-bracket-switch.png}
		\\
		(e) Matador \\
		(11, Valex 8)
	\end{tabular}
	\end{center}
	\caption{Miscellaneous interesting charts I discovered while browsing the giant spreadsheet.}
	\label{fig:hble-mention}
\end{figure*}

{\bf Sweet spot.}
Finally, in case it wasn't obvious in the tables, I'll point out that 9-15 is clearly the sweet spot of difficulties for technical stepcharts.

@@

Let's be honest: this isn't gonna be a paper trilogy.
Okay, with that said, here are some things that would be cool to implement in a fantasy universe with infinite free time.
(I have renamed this ``future'' work section accordingly.)

There are a few remaining cases the algorithm doesn't yet understand:
\begin{itemize}
	\item Doublesteps forced either by mine cues or by holds;
	\item Crossover and/or bracket jumps, not usually forced but often way less turny than the alternative;
	\item Forced footings across stream boundaries arising from bracket jumps or jump-footswitches.
\end{itemize}
For example, Figure~\ref{fig:hble-mention}(d) shows many sequential doublesteps, each forced by a mine cue,
but which the algorithm interprets as spins because the flipped stream length falls below \nflip.
Figure~\ref{fig:hble-mention}(e) shows an example of jump-footswitches
which the algorithm fails to count because it ignores the footing of jumps.

These patterns would all have to be identified heuristically.
Apart from that being more work than I wanted to do,
I also feel that adding too many heuristics to SIGBOVIK research
compromises the simple and innocent beauty of an implementation unbound by the demands of mainstream conferences.
